Multi-domain, multilingual translation with latent multi-task group dropout

### 2022-01-14 2022/acl/acl_latex.tex

Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi$ into the objective function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}. However, the operator $\operatorname{Top-k}$, which serves to define $\tilde{m}_l^d$ in Equation~\eqref{eq:top-k}, is not differentiable. Therefore, we approximate this function by the $\operatorname{Soft-Top-K}$ function defined as follows:
\begin{align}
\hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
                    0 \leqslant m_i \leqslant 1 \\
  \forall 0 \leqslant i \leqslant n_d\text{-}1 \\
        1^{T}.m = k
      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \label{eq:soft-top-k}
\end{align}
in which
\begin{align}
H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i). \nonumber 
\end{align}

In Appendix \ref{appendix:a}, we prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$. Furthermore, we also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and that its gradients can be computed using the implicit differentiation theorem. During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyper-parameter $\tau$ to $0$. The gradient of $D$ w.r.t $\Phi_l^d$ is computed using the chain rule as follows:
\begin{equation}
\frac{\partial D}{\partial \Phi_l^d} = \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}
\end{equation}
The gradient $\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.

We jointly train the Transformer parameters $\theta$ and the parameters of the variational distribution $\Phi$ using the following multi-task loss.\fyDone{This is repeated}
\begin{align}
\hspace{-0.em}
\mathcal{L}(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim MT^d(x)} \big[-\ell(x,y,d; \theta,\Phi)\big]
\end{align}

@@

We set the dropout probability to 0.1. We train the multi-domain Transformer model for 200k iterations with a batch size of 12k tokens. The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration. The \system{LaMGD} Transformer converged after 300k iterations with a batch size of 12k tokens. The convergence of \system{LaMGD} is controlled by its validation curve. Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline: this ensures that \system{LaMGD} and Adapter-based Transformer have the same amount of training iterations.

@@

The temperature parameter $\tau$ for the $\operatorname{Soft-Top-K}$ operator is gradually decreased from $0.5$ to $0.2$ according to the following policy:
\begin{align*}
\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \},
\end{align*}
in which $r=0.0001$.
While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we select an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical}. Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments
\subsubsection{Latent variables initialization}
We initialize the distribution of the latent variables uniformly. More precisely, we set $\Phi_l^d$, which generates the probability of the masks via the softmax activation function, to $0^{n_d}$.
\subsection{Datasets and metrics}
\subsubsection{Multidomain translation}
We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{See \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
    \cline{2-7} 
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
    \hline 
    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
    \hline
  \end{tabular}
  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$). \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
  }
\label{tab:Corpora-chap4}
\end{table*}
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks borrowing the multilingual translation datasets from past studies. More precisely, we used: 
\begin{itemize}
\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations from \citet{qi18when} between English and eight \emph{related languages}.
\item TED8-Diverse. The dataset consists of parallel sentences between English and eight \emph{diverse languages} as in \citet{Wang20balancing}.
\end{itemize}

@@

\subsection{Multidomain translation}
For these experiments, our main results are in Table~\ref{tab:mdmt}, where we observe that the \system{LaMGD} Transformer achieves a significant improvement (+2.78) over the generic \system{Transformer} system with zero extra parameters. Moreover, \system{LaMGD} Transformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is finetuned and contains approximately 25M additional parameters per domain. Variational mask learned from data by \system{LaMGD} also outperforms heuristic dropout mask \system{HMGD} by 0.5 in average.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[65m]} & \SB{40.3} & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
    \system{HMGD} Transformer   \hfill{\footnotesize[+0m]}  & \SB{40.4} & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & \SB{61.0} & \SB{53.1} & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & \SB{40.3} & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
    \hline
  \end{tabular}
  \caption{Multi-domain translation}
  \label{tab:mdmt}
\end{table*}

\subsubsection{Fuzzy domain separation}
\label{ssec:fuzzy}
For this experiment, we reuse proposal of \citet{Pham21revisiting}, who measure the efficiency of a multidomain NMT system exploiting the proximity between domains. It uses the same data as in the previous experiment; however, the domain \domain{law} is now randomly split into two pseudo-domains \domain{law$_1$} and \domain{law$_2$} of equal size. A truly multidomain system should be able to automatically detect the proximity between \domain{law$_1$} and \domain{law$_2$}, and there should be no significant difference between the performance of a system trained with the six original domains (including \domain{law}) or with the seven domains (including \domain{law} by \domain{law$_1$} and \domain{law$_2$}). The experiments in \citep{Pham21revisiting} reported a large gap between the two settings when using residual adapters. We replicated this setting and also report the results obtained with the \system{LaMGD} Transformer system in Table~\ref{tab:fuzzy}.

@@

\subsection{Multilingual translation}
Results for the multilingual experiments are in Table~\ref{tab:multilingual}.  The \system{LaMGD} Transformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively. Significant gains are observed for languages \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resource languages in our sets. However, \system{LaMGD} Transformer is outperformed by the multilingual \system{Transformer} and language Adapters for the O2M-diverse condition.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34.0}&38.3&8.3&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
    \hline
  \end{tabular}
  \caption{Multilingual Translation experiments. Boldface denotes significant gains over \system{Transformer} ($p=0.05$).\fyDone{What is the bold for ?}}
  \label{tab:multilingual}
\end{table*}

@@

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=0.7\textwidth]{multi_domain_heatmap.png}  
  \caption{Multidomain}
  \label{fig:md-heatmap}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=0.7\textwidth]{multilingual_heatmap.png}  
  \caption{Multilingual}
  \label{fig:ml-heatmap}
\end{subfigure}
\caption{Heatmap visualization of the similarities between dropout masks of domains(languages). Brighter colors correspond to higher similarity scores.}
\label{fig:heatmap}
\end{figure}

@@

We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d projection computed by PCA.
\begin{figure*}[htbp]
\begin{subfigure}{0.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=0.9\textwidth]{ted_related}  
  \caption{TED-Related}
  \label{fig:visual-related}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=0.85\textwidth]{ted_diverse}  
  \caption{TED-Diverse}
  \label{fig:visual-diverse}
\end{subfigure}
\caption{Visualization of languages according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:visual}
\end{figure*}

% \begin{figure*}[h!]
% \begin{subfigure}{0.5\textwidth}
%   \centering
%   % include first image
%   \includegraphics[width=0.9\textwidth]{phi_ted_related_o2m}  
%   \caption{TED-Related}
%   \label{fig:phi-related}
% \end{subfigure}
% \begin{subfigure}{0.5\textwidth}
%   \centering
%   % include second image
%   \includegraphics[width=0.85\textwidth]{phi_ted_diverse_o2m}  
%   \caption{TED-Diverse}
%   \label{fig:phi-diverse}
% \end{subfigure}
% \caption{$\Phi_L^{0}$ in O2M experiments}
% \label{fig:phi}
% \end{figure*}
\subsection{Ablation study}
We discuss here the choice of the hyper-parameters $k$, the number of activated nodes in each layer, and its impact on the sharing level between the tasks. Table ~\ref{tab:K} shows the variance of performance when the number of activated nodes is changed, and the sharing level between tasks decreases in consequence.\Todo{add sharing results. }\Todo{check caption consistency}
\begin{table}[h!]
  \centering
  \begin{tabular}{|c|c|c|} \hline
    $k$ & \domain{avg} & sharing rate \\ \hline 
    8 & 18.1 & 0.63\\
    10 & 19.15 & 0.73\\
    12 & 19.33 & 0.78\\
    14 & 19.44 & 0.88\\
    \hline
  \end{tabular}
  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$. These results correspond to the multilingual setting (o2m-related).}
  \label{tab:K}
\end{table}

>>>>

Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi$ into the objective function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}. However, the operator $\operatorname{Top-k}$, which serves to define $\tilde{m}_l^d$ in Equation~\eqref{eq:top-k}, is not differentiable. Therefore, we approximate this function by the $\operatorname{Soft-Top-K}$ function defined as follows:
\begin{align}\label{eq:soft-top-k}
\hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
                    0 \leqslant m_i \leqslant 1 \\
  \forall 0 \leqslant i \leqslant n_d\text{-}1 \\
        1^{T}.m = k
      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) 
\end{align}
in which
\begin{align}
H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i). \nonumber 
\end{align}

In Appendix \ref{appendix:a}, we prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$. Furthermore, we also provide the computation of $\hat{m}_l^d(\tau)$ and prove that $\hat{m}_l^d(\tau)$ is a differentiable function w.r.t $\Phi_l^d$ and that its gradients can be computed using the implicit differentiation theorem. During training, we approximate $\tilde{m}_l^d$ by $\hat{m}_l^d(\tau)$ by gradually decaying the hyper-parameter $\tau$ to $0.2$. The gradient of $D$ w.r.t $\Phi_l^d$ is computed using the chain rule as follows:
\begin{align*}
\frac{\partial D}{\partial \Phi_l^d} = \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} 
\end{align*}
The gradient $\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$ is computed via autograd algorithm while $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$ is computed via implicit differentiation, as explained in Appendix~\ref{appendix:a}.

We jointly train the Transformer parameters $\theta$ and the parameters of the variational distribution $\Phi$ using the following multi-task loss.\fyDone{This is repeated}
\begin{align*}
\hspace{-0.em}
\mathcal{L}(\theta,\Phi) = \sum_{d=1}^{n_d} \mathbb{E}_{x \sim \mathcal{D}_s^d,y \sim MT^d(x)} \big[-\ell(x,y,d; \theta,\Phi)\big]
\end{align*}

@@

We set the dropout probability to 0.1. We train the multidomain Transformer model for 200k iterations with a batch size of 12k tokens. The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration. The \system{LaMGD} Transformer converged after 300k iterations with the same batch size. The convergence of \system{LaMGD} is controlled by its validation curve. Finally, we plug adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.
%: this ensures that \system{LaMGD} and Adapter-based Transformer have the same amount of training iterations.

@@

The temperature parameter $\tau$ for the $\operatorname{Soft-Top-K}$ operator is gradually decreased from $0.5$ to $0.2$ according to the following policy:
\begin{align*}
\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \},
\end{align*}
in which $r=0.0001$.
While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we select an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical}. Finally, we set the weight of the entropy term to $0.0001$ in the training loss in every experiments.
%$\mathbb{H}(\text{softmax}(\Phi_l^d))$
\subsubsection{Latent variables initialization}
We initialize the distribution of the latent variables uniformly. More precisely, we set $\Phi_l^d$, which generates the probability of the masks via the softmax activation function, to $0^{n_d}$.
\subsection{Datasets and metrics}
\subsubsection{Multidomain translation}
We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{See \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
    \cline{2-7} 
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
    \hline 
    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
    \hline
  \end{tabular}
  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$).}
%\domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
\label{tab:Corpora-chap4}
\end{table*}
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks borrowing the multilingual translation datasets from past studies. More precisely, we used: 
\begin{itemize}
\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations from \citet{qi18when} between English and eight \emph{related languages}.
\item TED8-Diverse. The dataset consists of parallel sentences between English and eight \emph{diverse languages} as in \citet{Wang20balancing}.
\end{itemize}

@@

\subsection{Multidomain translation}
For these experiments, our main results are in Table~\ref{tab:mdmt}, where we observe that the \system{LaMGD} Transformer achieves a significant improvement (+2.78) over the generic \system{Transformer} system with zero extra parameters. Moreover, \system{LaMGD} Transformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is finetuned and contains approximately 25M additional parameters per domain. Variational mask learned from data by \system{LaMGD} also outperforms heuristic dropout mask \system{HMGD} by 0.5 in average.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
    \system{HMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.4 & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
    \hline
  \end{tabular}
  \caption{Multi-domain translation}
  \label{tab:mdmt}
\end{table*}

\subsubsection{Fuzzy domain separation}
\label{ssec:fuzzy}
For this experiment, we reuse proposal of \citet{Pham21revisiting}, who measure the efficiency of a multidomain NMT system exploiting the proximity between domains. It uses the same data as in the previous experiment; however, the domain \domain{law} is now randomly split into two pseudo-domains \domain{law$_1$} and \domain{law$_2$} of equal size. A truly multidomain system should be able to automatically detect the proximity between \domain{law$_1$} and \domain{law$_2$}, and there should be no significant difference between the performance of a system trained with the six original domains (including \domain{law}) or with the seven domains (including \domain{law} by \domain{law$_1$} and \domain{law$_2$}). \citet{Pham21revisiting} reported a large gap between the two settings when using residual adapters. We replicated this setting and report the results obtained with the \system{LaMGD} Transformer system in Table~\ref{tab:fuzzy}.

@@

\subsection{Multilingual translation}
Results for the multilingual experiments are in Table~\ref{tab:multilingual}.  The \system{LaMGD} Transformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively. Significant gains are observed for languages \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resource languages in our sets. However, \system{LaMGD} Transformer is outperformed by the multilingual \system{Transformer} and language Adapters for the O2M-diverse condition.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&34.0&38.3&8.3&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &10.2&4&\SB{13.3}&21.9&32.2&34.1&38.5&8.3&20.31 \\ 
    \system{LaMGD} Transformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
    \system{LaMGD} Transformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
    \hline
  \end{tabular}
  \caption{Multilingual Translation experiments. Boldface denotes significant gains over \system{Transformer} ($p=0.05$).\fyDone{What is the bold for ?}}
  \label{tab:multilingual}
\end{table*}

@@

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=0.7\textwidth]{multi_domain_heatmap.png}  
  \caption{Multidomain}
  \label{fig:md-heatmap}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=0.7\textwidth]{multilingual_heatmap.png}  
  \caption{Multilingual}
  \label{fig:ml-heatmap}
\end{subfigure}
\caption{Heatmap visualization of the similarities between dropout masks of domains(languages).}
\label{fig:heatmap}
\end{figure}

@@

We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d PCA projection.
\begin{figure*}[htbp]
\begin{subfigure}{0.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=0.9\textwidth]{ted_related}  
  \caption{TED-Related}
  \label{fig:visual-related}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=0.85\textwidth]{ted_diverse}  
  \caption{TED-Diverse}
  \label{fig:visual-diverse}
\end{subfigure}
\caption{Visualization of languages according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:visual}
\end{figure*}

% \begin{figure*}[h!]
% \begin{subfigure}{0.5\textwidth}
%   \centering
%   % include first image
%   \includegraphics[width=0.9\textwidth]{phi_ted_related_o2m}  
%   \caption{TED-Related}
%   \label{fig:phi-related}
% \end{subfigure}
% \begin{subfigure}{0.5\textwidth}
%   \centering
%   % include second image
%   \includegraphics[width=0.85\textwidth]{phi_ted_diverse_o2m}  
%   \caption{TED-Diverse}
%   \label{fig:phi-diverse}
% \end{subfigure}
% \caption{$\Phi_L^{0}$ in O2M experiments}
% \label{fig:phi}
% \end{figure*}
\subsection{Ablation study}
We discuss here the choice of the hyper-parameters $k$, the number of activated nodes in each layer, and its impact on the sharing level between the tasks. Table ~\ref{tab:K} shows the variance of performance when the number of activated nodes is changed, and the sharing level between tasks decreases in consequence.\Todo{add sharing results. }\Todo{check caption consistency}
\begin{table}[h!]
  \centering
  \begin{tabular}{|c|c|c|} \hline
    $k$ & \domain{avg} & sharing rate \\ \hline 
    8 & 18.1 & 0.63\\
    10 & 19.15 & 0.73\\
    12 & 19.33 & 0.78\\
    14 & 19.44 & 0.88\\
    \hline
  \end{tabular}
  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$ (o2m-related experiment).}
  \label{tab:K}
\end{table}
